The host galaxy of the short GRB~111117A at $z = 2.211$\thanks{Based on observations collected at the European Southern Observatory, Paranal, Chile, Program ID: 088.A-0051 and 091.D-0904.}

### 2017-06-01 GRB111117A.tex
Antonios comments
Closes #4

\begin{figure}
	\centering
	\includegraphics[width=9cm]{figures/GRB111117A_spec_obs.pdf}
	\caption{Imaging of the field of GRB~111117A with the X-shooter slit overlaid. Only one slit is shown, despite four epochs of spectroscopic observations because of the similarity in position angle. The image is the FORS2 $R$-band image and%, for which the photometry is shown in Fig. \ref{fig:SED}. 
	the blue asterisk indicates the GRB position as derived from the \emph{Chandra} observations in \citet{Sakamoto2013}. The errorcircle of the \emph{Chandra}-position is too small to be visible in the image.}
	\label{fig:spec_setup}
\end{figure}

Spectroscopic observations were carried out using the cross-dispersed echelle
spectrograph, VLT/X-shooter \citep{Vernet2011}, at four seperate epochs. The
burst was followed up 38 hours after the Burst Alert Telescope (BAT) trigger
under ESO programme 088.A-0051 (PI: Fynbo) and again later using a different ESO
program 091.D-0904 (PI: Hjorth). X-shooter can cover the wavelength range from
3000~\AA~to 24\,800~\AA~(21\,000~\AA~when the $K$-band blocking filter is used) across
three spectroscopic arms, simultaneously illuminated through the use of
dichroics. The bias-correction, flat-fielding, order tracing, wavelength
calibration, rectification, and flux calibration is then carried out using the
VLT/X-shooter pipeline version 2.8.4 \citep{Modigliani2010}. 
%Because the spectra are curved across each detector, a rectification algorithm
%is employed which introduces correlations between neighboring pixels. We select
%a pixel-scale of 0.2/0.2/0.6 \AA/pix for the UVB/VIS/NIR arm to minimize the
%degree of correlation while conserving the maximal resolution.
The observations are combined and extracted using scripts described in Selsing
et al. 2017 (in prep.) and available
online\footnote{\url{https://github.com/jselsing/XSGRB_reduction_scripts}}. The
signal-to-noise of the continuum in the near-infrared arm is too low to use the
optimal extraction algorithm \citep{Horne1986} and therefore the extraction is
carried out with a simple aperture.

An overview of the spectroscopic observations is given in Table~\ref{tab:spec_overview}, and the position of the slit on the target is shown in Fig.~\ref{fig:spec_setup}.

% \begin{figure}
% 	\centering
% 	\includegraphics[width=9cm]{figures/SEDspecphot.pdf}
% 	\caption{SED fit showning the best-fist SED to the derived photometry. The detection of \lya~ is predicted from the SED fit and confirmed by the spectroscopic observations. Overplotted in grey is a version of the observed spectrum binned by 30 \AA. The reason for the spectral gap at 10000 \AA is from the merging of the visual and near-infrared arm.}
% 	\label{fig:SED}
% \end{figure}

We determine a redshift of $z = 2.211$ from the simultaneous detection of
emission lines interpreted as \lya, \oii, \hb, \oiii$\lambda$5007, and \ha, with
\hb~detected at low significance ($\sim 3 \sigma$). \ha~is only visible in the
first epoch, due to the $K$-band blocking filter used for the remainder
observations. The nebular lines exhibit a spatial extent of $\sim$ 1\farc5 and
show significant velocity structure along the slit. A drop in the continuum to
the blue of the \lya~line further supports the inferred redshift.

@@

Using the luminosity of \ha, we can infer the star-formation rate (SFR) of a 
galaxy \citep{Kennicutt1998}. At the redshift of the GRB host, \ha~is observed at 
21\,000~\AA~where the night sky is very bright. In addition, several bright sky-lines 
intersect the line, making an accurate estimate of the \ha-flux difficult. 
We obtain a limit on the SFR by integrating the part of \ha~free of contamination and 
correcting for the missing flux using the line shape. After converting the 
\citet{Kennicutt1998} relation to a \citet{Chabrier2003} initial mass function using
\citet{Madau2014}, we derive a limit of $SFR > 7 M_\odot$~yr$^{-1}$. 
From the SED-fit (Sect. \ref{SED}), and the detection of \lya, the host is
constrained to contain very little or no dust, which is why we do not apply 
a dust-correction to the measured \ha~flux here.

@@

\begin{figure}
	\centering
	\includegraphics[width=9cm]{figures/OIII_img.pdf}
	\caption{2D-image of the \oiii$\lambda$5007 emission line. The location of a bright skyline is marked by the blue box. The location of the emission line is indicated with the red ellipse. Because the host is observed in nodding-mode, negative images of the emission line appear on both sides in the spatial direction.}
	\label{fig:line}
\end{figure}

@@

The multi-color spectral energy distribution (SED) is fit by
\citet{Bruzual2003} single stellar population models based on a
\citet{Chabrier2003} initial mass function in \emph{LePhare} \citep{Ilbert2006},
where the redshift is fixed to the spectroscopic value of $z=2.211$. The best
model is obtained with an unreddened galaxy template, and returns physical
parameters of absolute magnitude ($M_B=-22.0\pm0.1$\,mag), stellar mass
($\log(M_{\star}/M_\odot) = 9.9\pm0.2$), stellar population age ($\tau =
0.5_{-0.3}^{+0.5}$ Gyr) and star-formation rate
($SFR_{\mathrm{SED}}=11_{-4}^{+9} M_\odot\,\mathrm{yr}^{-1}$). %We show the SED
% fit in Fig. \ref{fig:SED}.

@@

We retrieved the automated data products provided by the \textit{Swift}-XRT GRB
repository\footnote{\url{http://www.swift.ac.uk/xrt\_products/00507901}}
\citep{Evans2009}. The X-ray afterglow light curve can be fit with a single power-law decay with
index $\alpha=1.27_{-0.10}^{+0.12}$. We performed a time-integrated spectral
analysis using data in photon counting (PC) mode in the widest time epoch where
the $0.3-1.5\,\mathrm{keV}$ to $1.5-10\,\mathrm{keV}$ hardness ratio is constant
(namely, from $t-T_0 = 205$~s to $t-T_0 = 203.5$~ks, for a total of 29.1~ks of
data) to prevent spectral changes that can affect the X-ray column density
determination. The obtained spectrum is well described by an absorbed power-law
model. The best-fit spectral parameters are a photon index of $2.1 \pm 0.4$ and
an intrinsic hydrogen column density $N_{\mathrm{H}}$ of $2.4_{-1.6}^{+2.4}
\times 10^{22}$~cm$^{-2}$ ($z=2.211$), assuming a Galactic $N_{\mathrm{H}}$ in
the burst direction of $4.1 \times 10^{20}$~cm$^{-2}$ \cite{Willingale2013}.

@@

As already pointed out \citep{Margutti2012, Sakamoto2013}, GRB~111117A securely 
belongs the short class of GRBs. Because the observed classification
indicators, $T_{90}$ and hardness ratio, do not depend strongly on redshift
\citep{Littlejohns2013a}, the updated redshift does not change this designation
significantly. The intrinsic luminosity increases, as visible in the X-ray light curve
in Fig. \ref{fig:sxray_lightcurve}, but it still is sub-luminous compared to
the majority of long GRBs. \citet{Bromberg2013} investigated the degree to which the
long and short population distributions overlap and quantified the certainty in
the class membership. GRB~111117A has $96_{-5}^{+3}$ percent probability of
being short. Compared to the other two sGRB candidates at high redshift, GRB060121
\citep{DeUgartePostigo2006, Levan2006} at $1.7 \lesssim z \lesssim 4.5$
($17_{-15}^{+14}$ per cent) and GRB090426 \citep{Antonelli2009, Levesque2010,
	Thone2011} at $z = 2.609$ ($10_{-10}^{+15}$ per cent), the certainty in class
membership for GRB~111117A is much higher.

\begin{figure}
	\centering
	\includegraphics[width=9cm]{figures/XLC_111117A_rest.pdf}
	\caption{Restframe XRT lightcurve of GRB~111117A, compared to the general population of XRT lightcurves of GRBs. The grey shared region is a compilation of long GRB lightcurves where the color represents density and the light blue is other short GRB lightcurves for which redshifts has been determined. Despite the remarkably high redshift, the luminosity is comparable to the bulk of the short burst population and still subluminous compared to the lGRB population.}
	\label{fig:sxray_lightcurve}
\end{figure}

@@

We plot the recalculated $N_\mathrm{H}$ in Fig. \ref{fig:NH_z} where we compare with the
distributions of complete samples of both long and short GRBs. The lGRB sample
is from \citet{Arcodia2016} and the sGRB sample is from \citet{DAvanzo2014a}.
17 of the 99 long bursts do not have redshifts and likewise 5 out of 16 for
the short sample. Bursts without redshifts have been excluded for both groups.
GRB~111117A occupies a unique position in Fig. \ref{fig:NH_z} with the highest
$N_H$ of all short burst. The short sample, excluding GRB~111117A, is located at
low redshifts ($z < 1$) and is found to populate a similar column density
environment to long GRBs at similar redshifts \citep{DAvanzo2014a}. The inferred
hydrogen column for GRB~111117A seems to follow the trend with increasing $N_H$
as a function of redshift as found for the long GRB afterglows \citep{Arcodia2016}.

\begin{figure}
	\centering
	\includegraphics[width=9cm]{figures/NH_z.pdf}
	\caption{Rest frame, X-ray derived hydrogen column densities for GRB~111117A compared to complete samples of both long and short populations. In red we show the sample of lGRBs from \citet{Arcodia2016}. The detections are replaced with contours for clarity and the limits are shown with arrows. In blue is the sample of sGRBs from \citet{DAvanzo2014a}. GRB~111117A is an outlier from the short sample, but seems to follow the distribution of long GRBs.}
	\label{fig:NH_z}
\end{figure}

@@

%From the clear host association, GRB~111117A does not belong to the hostless class of GRBs \citep{Berger2010a} and because the host exhibits emission lines indicative of a population of relatively young stars, 
As the majority of short GRBs \citep{Fong2013b}, the host of GRB~111117A is a
late-type galaxy and is entirely consistent in terms of stellar mass and stellar
age with the general population of short GRB hosts ($\left\langle M _*
\right\rangle = 10^{10.1} M_{\odot}$ and $\left\langle \tau _* \right\rangle =
0.3 $~Gyr) \citep{Leibler2010}. Being a late-type host, both the stellar mass
and sSFR are entirely within the range expected for the hosts of sGRBs
\citep{Behroozi2014}. The SFR is $\sim$1 order of magnitude higher than the
typical SFR for short GRB hosts galaxies \citep{Berger2014} and more similar to
the SFRs found in the hosts of long GRBs at a corresponding redshift
\citep{Kruhler2015}. The high SFR is partly a selection effect, because a less
star-forming galaxy would exhibit weaker emission lines, thus making the
redshift harder to determine. Additionally, it is natural to expect some
evolution in the hosts of sGRBs with redshift as illustrated in Sect.
\ref{restnH}.

@@

\begin{acknowledgements}
TK acknowledges support through the Sofja Kovalevskaja Award to P. Schady. CT acknowledges support from a Spanish National Research Grant of Excellence under project AYA 2014-58381-P and funding associated to a Ramón y Cajál fellowship under grant number RyC-2012-09984.
\end{acknowledgements}

@@

\newpage

@@

>>>>

\begin{figure}
	\centering
	\includegraphics[width=9cm]{figures/GRB111117A_spec_obs2.pdf}
	\caption{Imaging of the field of GRB~111117A with the X-shooter slit overlaid. Only one slit is shown, despite four epochs of spectroscopic observations because of the similarity in position angle. The image is the FORS2 $R$-band image for which the photometry is shown in Fig.~\ref{fig:SED}. 
	The blue asterisk indicates the GRB position as derived from the \emph{Chandra} observations in \citet{Sakamoto2013}. The insert shows the 2D-image of the \oiii$\lambda$5007 emission line. The location of a bright skyline is marked by the blue box. The location of the emission line is indicated with the red ellipse. Because the host is observed in nodding-mode, negative images of the emission line appear on both sides in the spatial direction.}
	\label{fig:spec_setup}
\end{figure}

Spectroscopic observations were carried out using the cross-dispersed echelle
spectrograph, VLT/X-shooter \citep{Vernet2011}, at four seperate epochs. The
burst was followed up 38 hours after the Burst Alert Telescope (BAT) trigger
under ESO programme 088.A-0051 (PI: Fynbo) and again later using a different ESO
program 091.D-0904 (PI: Hjorth). X-shooter can cover the wavelength range from
3000~\AA~to 24\,800~\AA~(21\,000~\AA~when the $K$-band blocking filter is used)
across three spectroscopic arms. The bias-correction, flat-fielding, order
tracing, wavelength calibration, rectification, and flux calibration is then
carried out using the VLT/X-shooter pipeline version 2.8.4
\citep{Modigliani2010}. %Because the spectra are curved across each detector, a
% rectification algorithm
%is employed which introduces correlations between neighboring pixels. We select
%a pixel-scale of 0.2/0.2/0.6 \AA/pix for the UVB/VIS/NIR arm to minimize the
%degree of correlation while conserving the maximal resolution.
The observations are combined and extracted using scripts described in Selsing
et al. 2017 (in prep.) and available
online\footnote{\url{https://github.com/jselsing/XSGRB_reduction_scripts}}. The
signal-to-noise of the continuum in the near-infrared arm is too low to use the
optimal extraction algorithm \citep{Horne1986} and therefore the extraction is
carried out with a simple aperture. An overview of the spectroscopic
observations is given in Table~\ref{tab:spec_overview}, and the position of the
slit on the target is shown in Fig.~\ref{fig:spec_setup}.We show the extracted
spectrum in Fig.~\ref{fig:SED}.

We determine a redshift of $z = 2.211$ from the simultaneous detection of
emission lines interpreted as \lya, \oii, \hb, \oiii$\lambda$5007, and \ha, with
\hb~detected at low significance ($\sim 3 \sigma$). We show \oiii$\lambda$5007
in the insert in Fig.~\ref{fig:spec_setup}.  \ha~is only visible in the first
epoch, due to the $K$-band blocking filter used for the remainder observations.
The nebular lines exhibit a spatial extent of $\sim$ 1\farc5 and show
significant velocity structure along the slit. A drop in the continuum to the
blue of the \lya~line further supports the inferred redshift.

@@

Using the luminosity of \ha, we can infer the star-formation rate (SFR) of a 
galaxy \citep{Kennicutt1998}. At the redshift of the GRB host, \ha~is observed at 
21\,000~\AA~where the night sky is very bright. In addition, several bright sky-lines 
intersect the line, making an accurate estimate of the \ha-flux difficult. 
We obtain a limit on the SFR by integrating the part of \ha~free of contamination and 
correcting for the missing flux using the line shape. After converting the 
\citet{Kennicutt1998} relation to a \citet{Chabrier2003} initial mass function using
\citet{Madau2014}, we derive a limit of $SFR > 7 M_\odot$~yr$^{-1}$. 
From the SED-fit (Sect.~\ref{SED}), and the detection of \lya, the host is
constrained to contain very little or no dust, which is why we do not apply 
a dust-correction to the measured \ha~flux here.

@@

%\begin{figure}
%	\centering
%	\includegraphics[width=9cm]{figures/OIII_img.pdf}
%	\caption{2D-image of the \oiii$\lambda$5007 emission line. The location of a bright skyline is marked by the blue box. The location of the emission line is indicated with the red ellipse. Because the host is observed in nodding-mode, negative images of the emission line appear on both sides in the spatial direction.}
%	\label{fig:line}
%\end{figure}

@@

The multi-color spectral energy distribution (SED) is fit by \citet{Bruzual2003}
single stellar population models based on a \citet{Chabrier2003} initial mass
function in \emph{LePhare} \citep{Ilbert2006}, where the redshift is fixed to
the spectroscopic value of $z=2.211$. The best model is obtained with an
unreddened galaxy template, and returns physical parameters of absolute
magnitude ($M_B=-22.0\pm0.1$\,mag), stellar mass ($\log(M_{\star}/M_\odot) =
9.9\pm0.2$), stellar population age ($\tau = 0.5_{-0.3}^{+0.5}$ Gyr) and
star-formation rate ($SFR_{\mathrm{SED}}=11_{-4}^{+9}
M_\odot\,\mathrm{yr}^{-1}$). We show the SED fit in Fig.~\ref{fig:SED}.

@@

We retrieved the automated data products provided by the \textit{Swift}-XRT GRB
repository\footnote{\url{http://www.swift.ac.uk/xrt\_products/00507901}}
\citep{Evans2009}. The X-ray afterglow light curve can be fit with a single power-law decay with
index $\alpha=1.27_{-0.10}^{+0.12}$. We performed a time-integrated spectral
analysis using data in photon counting (PC) mode in the widest time epoch where
the $0.3-1.5\,\mathrm{keV}$ to $1.5-10\,\mathrm{keV}$ hardness ratio is constant
(namely, from $t-T_0 = 205$~s to $t-T_0 = 203.5$~ks, for a total of 29.1~ks of
data) to prevent spectral changes that can affect the X-ray column density
determination. The obtained spectrum is well described by an absorbed power-law
model. The best-fit spectral parameters are a photon index of $2.1 \pm 0.4$ and
an intrinsic hydrogen column density $N_{\mathrm{H}}$ of $2.4_{-1.6}^{+2.4}
\times 10^{22}$~cm$^{-2}$ ($z=2.211$), assuming a Galactic $N_{\mathrm{H}}$ in
the burst direction of $4.1 \times 10^{20}$~cm$^{-2}$ \citep{Willingale2013}.

@@

As already pointed out \citep{Margutti2012, Sakamoto2013}, GRB~111117A securely 
belongs the short class of GRBs. Because the observed classification
indicators, $T_{90}$ and hardness ratio, do not depend strongly on redshift
\citep{Littlejohns2013a}, the updated redshift does not change this designation
significantly. 
%The intrinsic luminosity increases, as visible in the X-ray light curve
%in Fig. \ref{fig:sxray_lightcurve}, but it still is sub-luminous compared to
%the majority of long GRBs. 
\citet{Bromberg2013} investigated the degree to which the
long and short population distributions overlap and quantified the certainty in
the class membership. GRB~111117A has $96_{-5}^{+3}$ percent probability of
being short. Compared to the other two sGRB candidates at high redshift, GRB~060121
\citep{DeUgartePostigo2006, Levan2006} at $1.7 \lesssim z \lesssim 4.5$
($17_{-15}^{+14}$ per cent) and GRB~090426 \citep{Antonelli2009, Levesque2010,
	Thone2011} at $z = 2.609$ ($10_{-10}^{+15}$ per cent), the certainty in class
membership for GRB~111117A is much higher.

Additionally, \citet{Horvath2010} classifies both GRB~060121 and GRB~090426 as
intermediate duration bursts. This comes from both events having very soft
spectra, as compared to the hard ones typically seen in sGRBs. Intermediate
bursts are very clearly related in their properties to long GRBs
\citep{DeUgartePostigo2011}, so they are unlikely to come from compact object
mergers.

%\begin{figure}
%	\centering
%	\includegraphics[width=9cm]{figures/XLC_111117A_rest.pdf}
%	\caption{Restframe XRT lightcurve of GRB~111117A, compared to the general population of XRT lightcurves of GRBs. The grey shared region is a compilation of long GRB lightcurves where the color represents density and the light blue is other short GRB lightcurves for which redshifts has been determined. Despite the remarkably high redshift, the luminosity is comparable to the bulk of the short burst population and still subluminous compared to the lGRB population.}
%	\label{fig:sxray_lightcurve}
%\end{figure}

@@

We plot the recalculated $N_\mathrm{H}$ in Fig.~\ref{fig:NH_z} where we compare with the
distributions of complete samples of both long and short GRBs. The lGRB sample
is from \citet{Arcodia2016} and the sGRB sample is from \citet{DAvanzo2014a}.
17 of the 99 long bursts do not have redshifts and likewise 5 out of 16 for
the short sample. Bursts without redshifts have been excluded for both groups.
GRB~111117A occupies a unique position in Fig.~\ref{fig:NH_z} with the highest
$N_H$ of all short burst. The short sample, excluding GRB~111117A, is located at
low redshifts ($z < 1$) and is found to populate a similar column density
environment to long GRBs at similar redshifts \citep{DAvanzo2014a}. The inferred
hydrogen column for GRB~111117A seems to follow the trend with increasing $N_H$
as a function of redshift as found for the long GRB afterglows \citep{Arcodia2016}.

\begin{figure}
	\centering
	\includegraphics[width=9cm]{figures/NH_z.pdf}
	\caption{Rest frame, X-ray derived hydrogen column densities for GRB~111117A compared to complete samples of both long and short populations. The detections are replaced with contours for clarity and the limits are shown with arrows.}
	\label{fig:NH_z}
\end{figure}

@@

%From the clear host association, GRB~111117A does not belong to the hostless class of GRBs \citep{Berger2010a} and because the host exhibits emission lines indicative of a population of relatively young stars, 
As the majority of short GRBs \citep{Fong2013b}, the host of GRB~111117A is a
late-type galaxy and is entirely consistent in terms of stellar mass and stellar
age with the general population of short GRB hosts ($\left\langle M _*
\right\rangle = 10^{10.1} M_{\odot}$ and $\left\langle \tau _* \right\rangle =
0.3 $~Gyr) \citep{Leibler2010}. Being a late-type host, both the stellar mass
and sSFR are entirely within the range expected for the hosts of sGRBs
\citep{Behroozi2014}. The SFR is $\sim$1 order of magnitude higher than the
typical SFR for short GRB hosts galaxies \citep{Berger2014} and more similar to
the SFRs found in the hosts of long GRBs at a corresponding redshift
\citep{Kruhler2015}. The high SFR is partly a selection effect, because a less
star-forming galaxy would exhibit weaker emission lines, thus making the
redshift harder to determine. Additionally, it is natural to expect some
evolution in the hosts of sGRBs with redshift as illustrated in Sect.~\ref{restnH}.

@@

\begin{acknowledgements}
TK acknowledges support through the Sofja Kovalevskaja Award to P. Schady. 
CT acknowledges support from a Spanish National Research Grant of Excellence under project AYA 2014-58381-P and funding associated to a Ramón y Cajál fellowship under grant number RyC-2012-09984.
AdUP acknowledges support from a Ramón y Cajal fellowship, a BBVA Foundation Grant for Researchers and Cultural Creators,and the Spanish Ministry of Economy and Competitiveness through project AYA2014-58381-P.
Partly based on observations made with the Gran Telescopio Canarias (GTC).

\end{acknowledgements}

@@

\newpage
%\appendix

@@

 \begin{figure*}
 	\centering
 	\includegraphics[width=16cm]{figures/SEDspecphot.pdf}
 	\caption{SED fit showning the best-fist SED to the derived photometry. The detection of \lya~ is predicted from the SED fit and confirmed by the spectroscopic observations. Overplotted in grey is the observed spectrum. The reason for the spectral gaps at 5500 \AA~and 10000 \AA~is from the merging of the arms.}
 	\label{fig:SED}
 \end{figure*}

### 2017-06-01 tables/phot_overview.tex
Antonios comments
Closes #4

	\centering
	\caption{Overview of the photometric observations. \label{tab:phot_overview}}
	\begin{tabular}{ccccccc}
		\hline\hline
{Obs. Date} &  Exptime & Telescope/Instrument & Filter & Airmass & Image Quality & Host Brightness\tablefootmark{a}  \\ [1.5pt]
        \hline
{} & {ks} &    & {} & & (arcsec)  & (mag$_{\mathrm{AB}}$)  \\ [1.5pt]
		\hline
2013-08-30T07:43 & 1.45 & VLT/FORS2 & $g$ & 1.55 & 0.99 & $24.08\pm 0.09$ \\
2011-11-17T20:07 & 0.80 & GTC/OSIRIS & $g$ & 1.15 & 1.67 & $24.13\pm 0.09$ \\
2011-11-17T20:07 & 1.20 & GTC/OSIRIS & $r$ & 1.11 & 1.50 & $23.93\pm 0.08$ \\        
2013-07-17T08:37 & 1.45 & VLT/FORS2 & $R$ & 1.56 & 0.74 & $23.95\pm 0.06$ \\   
2011-11-28T21:10 & 3.60 & TNG/LRS & $R$ & 1.01 & 1.08 & $23.96\pm 0.13$ \\           
2011-11-17T20:07 & 0.36 & GTC/OSIRIS & $i$ & 1.08 & 1.50 & $23.89\pm 0.23$ \\   
2013-08-03T09:23 & 1.35 & VLT/FORS2 & $I$ & 1.54 & 0.93 & $24.22\pm 0.15$ \\           
2011-11-28T06:14 & 1.80 & Gemini/GMOS-N & $z$ & 1.01 & 0.84 & $24.24\pm 0.47$ \\  
2013-07-13T09:33 & 1.08 & VLT/FORS2 & $z$ & 1.49 & 0.63 & $23.76\pm 0.21$ \\             
2013-06-24T09:14 & 1.98 & VLT/HAWK-I & $J$ & 1.70 & 0.63 & $23.13\pm 0.18$ \\        
2013-06-27T09:21 & 1.68 & VLT/HAWK-I & $H$ & 1.63 & 0.91 & $22.94\pm 0.29$ \\   
2013-06-28T09:14 & 1.92 & VLT/HAWK-I & $K_\mathrm{s}$ & 1.65 & 0.76 & $23.07\pm 0.32$ \\   
\hline\noalign{\smallskip}

>>>>

	\centering
	\caption{Overview of the photometric observations. \label{tab:phot_overview}}
	\begin{tabular}{ccccccc}
		\hline\hline
{Obs. Date} &  Exptime & Telescope/Instrument & Filter & Airmass & Image Quality & Host Brightness\tablefootmark{a}  \\ [1.5pt]
        \hline
{} & {ks} &    & {} & & (arcsec)  & (mag$_{\mathrm{AB}}$)  \\ [1.5pt]
		\hline
2013-08-30T07:43 & 1.45 & VLT/FORS2 & $g$ & 1.55 & 0.99 & $24.08\pm 0.09$ \\
2011-11-17T20:07 & 0.80 & GTC/OSIRIS & $g$ & 1.15 & 1.67 & $24.13\pm 0.09$ \\
2011-11-17T20:07 & 1.20 & GTC/OSIRIS & $r$ & 1.11 & 1.50 & $23.93\pm 0.08$ \\        
2013-07-17T08:37 & 1.45 & VLT/FORS2 & $R$ & 1.56 & 0.74 & $23.95\pm 0.06$ \\   
2011-11-28T21:10 & 3.60 & TNG/DOLORES & $R$ & 1.01 & 1.08 & $23.96\pm 0.13$ \\           
2011-11-17T20:07 & 0.36 & GTC/OSIRIS & $i$ & 1.08 & 1.50 & $23.89\pm 0.23$ \\   
2013-08-03T09:23 & 1.35 & VLT/FORS2 & $I$ & 1.54 & 0.93 & $24.22\pm 0.15$ \\           
2011-11-28T06:14 & 1.80 & Gemini/GMOS-N & $z$ & 1.01 & 0.84 & $24.24\pm 0.47$ \\  
2013-07-13T09:33 & 1.08 & VLT/FORS2 & $z$ & 1.49 & 0.63 & $23.76\pm 0.21$ \\             
2013-06-24T09:14 & 1.98 & VLT/HAWK-I & $J$ & 1.70 & 0.63 & $23.13\pm 0.18$ \\        
2013-06-27T09:21 & 1.68 & VLT/HAWK-I & $H$ & 1.63 & 0.91 & $22.94\pm 0.29$ \\   
2013-06-28T09:14 & 1.92 & VLT/HAWK-I & $K_\mathrm{s}$ & 1.65 & 0.76 & $23.07\pm 0.32$ \\   
\hline\noalign{\smallskip}
